Negative Selection in the 20th Century

### 2020-05-14 why-negative-selection.Rmd
weight data using GHS 2006.

Despite this, effect sizes tend to be smaller for children regressions. This
could be caused by ascertainment bias in the UK Biobank sample -- e.g.,
if respondents themselves are a more selected sample than the respondents' 
parents.

@@

Different polygenic scores are correlated. Table \@ref(tab:pgs-corrs) shows the top
correlations in the sample. Because of this, bivariate correlations between
PGS and number of children might be driven by other genetic scores. To explore which polygenic scores are driving negative selection, we run a single
omnibus regression of *number of children* on all the PGS. We exclude EA2, 
waist-hip ratio, WC (XXX what is it?) and "Hip combined" since they are 
highly correlated with other scores, which could make our estimates unstable. Figure \@ref(fig:plot-together) shows the results. Interestingly, several PGS remain independently significant, although
effect sizes are reduced.

>>>>

```{r res-weighted}

drake::loadd(res_weighted)
res_weighted_all <- res_weighted %>% 
      left_join(
        res_all %>% 
              filter(dep.var == "N children", reg.type == "raw") %>% 
              select(term, estimate),
        by = "term",
        suffix = c("_weighted", "_unweighted")
      ) %>% 
      mutate(
        prop_wtd_unwtd = estimate_weighted/estimate_unweighted
      )

mean_prop_wtd_unwtd <- res_weighted_all %>% 
      filter(sign(estimate_weighted) == sign(estimate_unweighted)) %>% 
      summarize(mean = mean(prop_wtd_unwtd)) %>% 
      pull(mean)

# Accounting for ascertainment bias

Effect sizes tend to be smaller for children regressions. This
could be caused by ascertainment bias in the UK Biobank sample -- e.g.,
if respondents themselves are a more selected sample than the respondents' 
parents. To check this, we weighted UK Biobank participants by age of leaving
full time education. Our sample weights are based on the 2006 General Household
Survey, calculating proportions within age and sex cells among White British
respondents. We then rerun the basic *children* regressions.

Figure \@ref(fig:plot-weighted) shows the results, with unweighted estimates
plotted for comparison. Weighting increases effect sizes on average by 
`r 100 * (mean_prop_wtd_unwtd - 1)` per cent. These weights are only basic 
attempts to correct for ascertainment bias. Effect sizes might be increased 
further by more precise weighting.

```{r plot-weighted, fig.cap = "Effects of polygenic scores on number of children, regressions weighted by education levels within age categories"}

n_regs <- as.double(nrow(res_weighted_all))

res_weighted_all %>% 
      mutate(
        term    = pretty_names(term),
        term    = fct_reorder(term, estimate_weighted, order_abs)
      ) %>% 
      pivot_longer(cols = c(estimate_weighted, estimate_unweighted), 
        names_to = "Regression", values_to = "estimate") %>% 
      mutate(
        Regression = sub("estimate_", "", Regression)
      ) %>% 
      ggplot(aes(x = term, y = estimate, fill = Regression,
        colour = p.value < 0.05/{{n_regs}})) +
      my_hline + 
      geom_point(size = 1.5, stroke = .8, shape = 21) + 
      labs(y = "Effect size", x = "") +
      coord_flip() +
      scale_fill_manual(values = 
        c("unweighted" = "transparent", "weighted" = "darkgreen")) +
      scale_colour_manual(values = pval_palette)

@@

Different polygenic scores are correlated. Table \@ref(tab:pgs-corrs) shows the top
correlations in the sample. Because of this, bivariate correlations between
PGS and number of children might be driven by other genetic scores. To explore which polygenic scores are driving negative selection, we run a single
omnibus regression of *number of children* on all the PGS. We exclude EA2, 
waist-hip ratio, waist-circumference, and "Hip combined" since they are 
highly correlated with other scores, which could make our estimates unstable. Figure \@ref(fig:plot-together) shows the results. Interestingly, several PGS remain independently significant, although
effect sizes are reduced.
